---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The data is Seattle Public Library checkouts from April 2005 to January 2023 from one of my favorite sci-fi authors, Margarette Peterson Haddix. I want to analyze how the number of checkouts increases or decreases over time because I want to see if there were certain years where Haddix had a lot of popularity. This is the same reason for wanting to find the year with the most Haddix book checkouts. I also want to see the number of books published over time to see which years was Haddix the most and least productive in writing. I also want to
see which books had the most and least checkouts overall, so I find out which books are the most and least popular within those 18 years of publishing. 

### Summary Information

1. Total number of books documented from checkouts from Haddix from April 2005 to January 2023 - 3604
2. Total number of years Haddix books were published from April 2005 to January 2023 - 24
3. The most number of checkouts per book - 25
4.The year with the most number of Haddix book checkouts - 2011
5. The year with the least number of Haddix book checkouts - 2023
6. The average number of Haddix book checkouts per year - 683

### The Dataset

- Who collected/published the data?  
The Seattle Public Library (SPL) collected and published this dataset.

- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
  There are 3604 observations and 12 features in the dataset. They are UsageClass (Physical or Digital), CheckoutType (what system was used for checkout), BookType (book, ebook, audiobook), CheckoutYear & CheckoutMonth, Checkouts (Number of Checkouts), Title, ISBN, Creator, Subjects, Publisher, and Publication Year.
   
- How was the data collected or generated?  
  SPL collects both physical and digital checkouts. Physical checkouts were collected by Legrady Artwork data archives, Horizon ILS, and other sources. Digital checkouts were collected by Overdrive, Hoopla, and RGDigital data.
   
- Why was the data collected?  
  The Seattle Public Library (SPL) provided this dataset of checkouts to the public to not only keep a record of checkouts, but also provide a deeper understanding of what Seattle readers like and on any given timeframe. 

- What, if any, ethical questions do you need to consider when working with
this data?  
  Whether or not the data includes personal information about those who checked out books. 

- What are possible limitations or problems with this data?   (at least 200 words)
  Checkouts are limited to the Seattle Public Library, so it is only representative of SPL readers who check out books. Just because someone checked out a book from the library does not mean that they show interest in it; books may be checked out for many reasons other than showing interest, such as having a gift, or having a required book for school. Also, the data is updated on the sixth day of every month, so the data is not as current as data updated every day. The data only counts the first checkout, so any renews, like number of times the book was renewed, are not listed.

### First Trends Over Time Chart

This shows the relationship between sci-fi/juvenile fiction checkouts and year (Apr 2005-Jan 2023), so one can see how those two genres compare over the years in terms of popularity. Juvenile peaked in checkouts in 2011, while Sci-Fi peaked in checkouts in 2012.

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart1.R")}
plot(chart1)
```

### Second Trends Over Time Chart

This shows the number of books published under Sci-Fi and Juvenile Fiction for each publication year (Apr 2005-Jan 2023). It shows that Haddix published her books consistently from 2012 to 2019 for both genres.


``````{r, echo = FALSE, code = readLines("chart2.R")}
plot(chart2)
```

### Your Choice


``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```


